\subsection{Hao-Orlin}
Our goal in this section is to find a minimum capacity non-trivial cut in $D = (N, A)$.

\begin{example}{}{}
    Suppose we are given the following directed graph:

    \begin{minipage}{\textwidth}
        \centering
        \begin{tikzpicture}
            \node[vertex] (a) at (0,0) {a};
            \node[vertex] (b) at (0,3) {b};
            \node[vertex] (c) at (3,3) {c};
            \node[vertex] (d) at (3,0) {d};

            \begin{scope}[every node/.style={fill = white, ellipse}]
                \draw[edge] (a) to node {$8$} (b);
                \draw[edge] (b) to node {$5$} (c);
                \draw[edge] (c) to node {$7$} (d);
                \draw[edge] (d) to node {$4$} (a);
                \draw[edge] (a) to node {$10$} (c);
            \end{scope}
        \end{tikzpicture}
    \end{minipage}

    The cut $\delta(S)$ where $S = \{a,b\}$ has capacity $c(\delta(S)) = 10 + 5$ (We ignore incoming arcs)

    The cut $\delta(T)$ where $T = \{c,d\}$ has capacity $c(\delta(T)) = 4$
\end{example}

\underline{Naive Algorithm}: We can pick all possible pairs of nodes $s,t$ and run a maximum flow algorithm on each pair. 
There are $\mathcal{O}(|N|^2)$ such pairs, so we'll have to do this $\mathcal{O}(|N|^2)$ times.

\underline{But}: we can be more efficient than that. Suppose we pick any $s \in N$. 
$s$ is either in the set that makes the global minimum cut, or it isn't. 
So, we just need to find all $s,t$-cuts and $t,s$-cuts and output the minimum.
This only takes $2|N| - 1$ iterations of a maximum flow algorithm.

We can do better still: Suppose we had an algorithm that gave the minimum cut containing $s$ (efficiently).
Then, we only need to run this algorithm for every $s \in N$ to find the global minimum cut.

We'll call this the \underline{minimum $s$-cut} problem:
\begin{definition}{$s$-cut}{}
    Given $s \in N$, an \underline{$s$-cut} has the form $\delta(S)$ where $s \in S$ and $S \neq N$
\end{definition}

The goal of Hao-Orlin is to solve the minimum $s$-cut problem quickly.
It is a modification of the push-relabel algorithm

\subsubsection{Preparatory Work}
We'll need to tinker with the push-relabel algorithm a bit in order to make the algorithm do what we want.
Here, we'll make some definitions and lay the groundwork for Hao-Orlin.

\begin{definition}{$X,t$-cut}{}
    Given $X \subseteq N$ and $t \not\in X$, an \underline{$X,t$-cut} has the form $\delta(S)$, where $X \subseteq S, t \not\in S$

    \begin{minipage}{\textwidth}
        \centering
        \begin{tikzpicture}[node distance = 2cm]
          \node[vertex, minimum size = 1.5cm] (X) at (1,0) {$X$};
          \node[vertex] (t) at (5,0) {$t$};
          \node[vertex, minimum size = 2.5cm, dotted][label=below left:{$S$}] (S) at (1,0) {};
    
          \draw[edge] (S) to (4, 1); 
          \draw[edge] (S) to (4, 0); 
          \draw[edge] (S) to (4, -1); 
        \end{tikzpicture}
        \captionof*{figure}{An $X,t$-cut}
      \end{minipage}
\end{definition}

Suppose we are given the following algorithm:

\IncMargin{1em}
\begin{algorithm}[H]\label{alg:generic-s-cut}  
    \nl Set $X = \{s\}$
    
    \nl \While{$X \neq N$}{
        \begin{enumerate}[label=(\alph*)]
            \item Pick $t \not\in X$, find a minimum $X,t$-cut
            \item Add $t$ to $X$
        \end{enumerate}
    }
    
    \nl Output minimum cut over all cuts found. 

    \caption{Generic Algorithm for minimum $s$-cut}
\end{algorithm}

We claim that this helps us find the minimum $s$-cut.

Let $\delta(S)$ be a minimum $s$-cut.
Consider the first time we pick some $t \not\in S$ in \Cref{alg:generic-s-cut}.
The algorithm gives us a minimum $X,t$-cut $\delta(S^*)$ and since we started with $X = \{s\}$, this is also an $s$-cut.
But, $\delta(S)$ is also an $X,t$-cut.
So, both cuts must have the same capacity.

In fact, Hao-Orlin will build on this algorithm.
But first, we need to modify a couple more aspects of the push-relabel algorithm.

\begin{definition}{$X$-preflow}{}
    For $X \subseteq N$, an \underline{$X$-preflow} is a flow $x$ where $e_x(v) \geq 0$ whenever $v \not\in X$
\end{definition}

We will keep our definition of \underline{height} the same.
But, we will need to redefine compatibility.

\begin{definition}{Compatible}{}
    Heights $h$ are \underline{compatible} with an $X$-preflow if:
    \begin{enumerate}
        \item $h(v) = |N|$ for all $v \in X$
        \item $h(t) \leq |X| - 1$
        \item $h(v) \geq h(u) - 1$ for all $uv \in D'$
    \end{enumerate}
\end{definition}

And we'll introduce two new concepts:
\begin{definition}{Level, Cut-Level}{}
    A \underline{level} $k$ consists of all nodes with height $k$, denoted $H(k)$

    A \underline{cut-level} is a level $k$ where no arc goes from $H(k)$ to $H(k-1)$ in $D'$
\end{definition}

These two definitions will allow us to rewrite the following lemma using terminology that will show up in our final algorithm:

\begin{lemma}{}{}
    If $\delta(S)$ is an $X,t$-cut with $\delta_{D'}(S) = \emptyset$ and $e_x(v) = 0$ for all $v \in N\setminus(S \cup \{t\})$, then $\delta(S)$ is a minimum $X,t$-cut.
\end{lemma}

Using the definitions of level and cut-level, we can rewrite this lemma to the following:
\begin{corollary}{}{cut-level-min-cut}
    If $\ell$ is a cut level and $e_x(v) = 0$ for all $v$ with $h(v) < \ell$, except $t$, then $\{v\:|\:h(v) \geq \ell\}$ is a minimum $X,t$-cut.
\end{corollary}

Let's prove the lemma:
\begin{proof}
    We want to show that $x$ is maximum $X,t$-flow.

    First, let's establish some bounds. By \Cref{lem:ff-correctness}, we can bound the net outflow of $S$:
    \begin{equation*}
        x(\delta(S)) - x(\delta(\overline{S})) \leq c(\delta(S))
    \end{equation*}

    This flow goes to $N\setminus S$ so:
    \begin{equation*}
        \sum_{v \in N \setminus S} e_x(v) = x(\delta(S)) - x(\delta(\overline{S})) \leq c(\delta(S))
    \end{equation*}
    
    Finally, the excess at $t$ is at most the inflow from all nodes in $N \setminus S$, so:
    \begin{equation}\label{eq:xt-flow-bound}
        e(t) \leq \sum_{v \in N \setminus S} e_x(v) = x(\delta(S)) - x(\delta(\overline{S})) \leq c(\delta(S))
    \end{equation}

    We're done if we can show that the bounds in \Cref{eq:xt-flow-bound} are tight.
    This shows that $x$ attains the upper bound and $t$ obtains all the excess.

    Using the assumption $\delta_{D'}(S) = \emptyset$, it must be that:
    \begin{equation*}
        x(\delta(S)) - x(\delta(\overline{S})) = c(\delta(S))
    \end{equation*}

    Using the assumption that $e_x(v) = 0$ for all $v \in N\setminus(S \cup \{t\})$:
    \begin{equation*}
        e(t) = \sum_{v \in N \setminus S} e_x(v)
    \end{equation*}

    So, $\delta(S)$ is a minimum $X,t$-cut.
\end{proof}

\subsubsection{The Algorithm}
\underline{Idea}: We want to use \Cref{corol:cut-level-min-cut}. To that end, we'll keep track of a cut-level $\ell$ and aim to get rid of excess on nodes below $\ell$ while running push-relabel.

\begin{note}
    We'll keep an \underline{invariant}: Non-empty levels will be kept consecutive (This will be useful later)
\end{note}

\begin{algorithm}[H]\label{alg:hao-orlin}
    \nl \textbf{Initialization:}
    \begin{itemize}
        \item Set $X = \{s\}$ ($s$ arbitrary)
        \item Choose $t \in N\setminus X$
        \item Set $h(s) = |N|$ and $h(v) = 0$ otherwise
        \item Set the cut level $\ell$ to be $|N| - 1$
        \item Send as much flow out of $X$ as possible
    \end{itemize}
    
    \nl \While{$X \neq N$}{
        \begin{enumerate}
            \item Run push-relabel with the following exceptions:
            \begin{enumerate}
                \item\label{item:hao-orlin-exception-1} Only select nodes $v$ such that $e(v) > 0$ and $h(v) < \ell$
                \item\label{item:hao-orlin-exception-2} If we want to relabel $v$ and $v$ is the only node with height $h(v)$.
                
                Do not relabel. Instead, set $\ell = h(v)$
                \item\label{item:hao-orlin-exception-3} If we want to relabel $v$ to $\ell$, then relabel and reset $\ell = |N| - 1$
            \end{enumerate}

            \item When no node satisfies $e(v) > 0$ and $h(v) < \ell$:
            \begin{enumerate}
                \item Store the cut $\{v \::\: h(v) \geq \ell \}$ (This is a min $X,t$-cut)
                \item Add $t$ to $X$
                \item Set $h(t) = |N|$, send as much flow out of $X$ as possible, and pick the new $t$ to be the node with the lowest height (This step maintains comp)
                \item Reset $\ell = |N| - 1$
            \end{enumerate}  
        \end{enumerate}
    }
    
    \nl Output minimum cut over all cuts found. 

    \caption{Hao-Orlin Algorithm}
\end{algorithm}
Some notes on the exceptions:
\begin{enumerate}
    \item Step \ref{item:hao-orlin-exception-1} ensures that we can apply \Cref{corol:cut-level-min-cut}
    \item Step \ref{item:hao-orlin-exception-2} maintains the invariant
    \item Step \ref{item:hao-orlin-exception-3} is used to maintain the cut level. For example, suppose there is an arc $uv$ where $h(u) = h(v) = \ell - 1$, and we want to relabel $u$ to $\ell$. This would violate the condition as $uv$ is an arc going across $H(\ell)$ and $H(\ell - 1)$.
\end{enumerate}
    

\subsubsection{Correctness}